---
title: "Tidy Simulation and Power Analyses"
author: "Brandon LeBeau"
output: 
  revealjs::revealjs_presentation:
    theme: simple
    highlight: pygments
    transition: fade
    mathjax: null
    slide_level: 1
    includes:
      in_header: custom.css
    reveal_options:
      slideNumber: true
      conrols: false
editor_options: 
  chunk_output_type: console
---

```{r echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(simglm)
```


# `simglm` package for simulation and power analyses

+ Simulates a variety of regression models
   - Continuous, binary, or count outcomes
   - With or without random effects
   - Longitudinal mixed models
+ Flexible model specification
   - Missing data
   - Correlated data
   - Population vs Sample specific models
+ Replication built into package
+ Parallel processing using `future`
+ Current CRAN version: 0.8.8


# Modular Approach

- **Simulation**
    + `simulate_*`: These functions perform a simulation operation
    + `correlate_*`: These correlate simulation based elements.
    + `generate_*`: These generate based on the simulation steps
- **Power/Model**
    + `model_fit()`: Fit a statistical model to simulated data
    + `extract_coefficients()`: Extract model coefficients
    + `replicate_simulation()`: Replicate simulation and/or model fitting
    + `compute_statistics()`: Compute statistics to summarize replication
    
    
# Simulation Example

```{r }
set.seed(321) 

sim_arguments <- list(
  formula = y ~ 1 + weight + age + sex,
  fixed = list(weight = list(var_type = 'continuous', mean = 180, sd = 30),
               age = list(var_type = 'ordinal', levels = 30:60),
               sex = list(var_type = 'factor', levels = c('male', 'female'))),
  error = list(variance = 25),
  sample_size = 1000,
  reg_weights = c(2, 0.3, -0.1, 0.5)
)

simulate_fixed(data = NULL, sim_arguments) %>%
  simulate_error(sim_arguments) %>%
  generate_response(sim_arguments) %>% 
  head()
```

# Model Fitting

```{r }
simulate_fixed(data = NULL, sim_arguments) %>%
  simulate_error(sim_arguments) %>%
  generate_response(sim_arguments) %>% 
  model_fit(sim_arguments) %>%
  extract_coefficients()
```

# Power Analysis - simulation arguments
+ New Simulation arguments for power

```{r}
sim_arguments <- list(
  formula = y ~ 1 + weight + age + sex,
  sample_size = 50,
  model_fit = list(formula = y ~ 1 + weight + age + sex,
                   model_function = 'lm'),
  replications = 1000,
  power = list(
    dist = 'qt',
    alpha = .02,
    opts = list(df = 49)
  ),
  extract_coefficients = TRUE
)
```


```{r echo = FALSE}
set.seed(321) 

sim_arguments <- list(
  formula = y ~ 1 + weight + age + sex,
  fixed = list(weight = list(var_type = 'continuous', mean = 180, sd = 30),
               age = list(var_type = 'ordinal', levels = 30:60),
               sex = list(var_type = 'factor', levels = c('male', 'female'))),
  error = list(variance = 25),
  sample_size = 50,
  reg_weights = c(2, 0.3, -0.1, 0.5),
  model_fit = list(formula = y ~ 1 + weight + age + sex,
                   model_function = 'lm'),
  replications = 1000,
  power = list(
    dist = 'qt',
    alpha = .02,
    opts = list(df = 49)
  ),
  extract_coefficients = TRUE
)
```

# Power Analysis - Replications

```{r}
library(future)
plan(multisession)

replicated_simulations <- replicate_simulation(sim_arguments)
```

+ `replicate_simulation()` is basically:

```{r eval = FALSE}
simulate_fixed(data = NULL, sim_arguments) %>%
  simulate_error(sim_arguments) %>%
  generate_response(sim_arguments) %>% 
  model_fit(sim_arguments) %>%
  extract_coefficients()
```


# Power Analysis - Compute Statistics
```{r}
compute_statistics(replicated_simulations, sim_arguments,
                   power = TRUE, 
                   type_1_error = FALSE,
                   precision = TRUE)
```

# Vary Simulation Arguments
+ New arguments for varying arguments.

```{r}
sim_arguments <- list(
  formula = y ~ 1 + weight + age + sex,
  vary_arguments = list(
    sample_size = c(50, 100, 200, 1000)
  )
)
```

```{r echo = FALSE}
set.seed(321) 

sim_arguments <- list(
  formula = y ~ 1 + weight + age + sex,
  fixed = list(weight = list(var_type = 'continuous', mean = 180, sd = 30),
               age = list(var_type = 'ordinal', levels = 30:60),
               sex = list(var_type = 'factor', levels = c('male', 'female'))),
  error = list(variance = 25),
  reg_weights = c(2, 0.3, -0.1, 0.5),
  model_fit = list(formula = y ~ 1 + weight + age + sex,
                   model_function = 'lm'),
  replications = 1000,
  power = list(
    dist = 'qt',
    alpha = .02,
    opts = list(df = 49)
  ),
  extract_coefficients = TRUE,
  vary_arguments = list(
    sample_size = c(50, 100, 200, 1000)
  )
)
```

# Power - Replications
```{r}
library(future)
plan(multisession)

replicated_simulations <- replicate_simulation(sim_arguments)
```

# Power - Compute Statistics
```{r}
power_est <- compute_statistics(replicated_simulations, sim_arguments,
                   power = TRUE, 
                   type_1_error = FALSE,
                   precision = TRUE)

head(power_est)
```

# Power Visualization
```{r echo = FALSE}
library(ggplot2)

ggplot(power_est, aes(x = as.numeric(sample_size), y = power, color = term)) + 
  geom_line(aes(group = term), size = 1.25) + 
  geom_point(size = 4) + 
  theme_bw(base_size = 18) + 
  scale_x_continuous("Sample Size", 
                     breaks = c(50, 100, 200, 1000)) + 
  ylab("Empirical Power")
```



# More Details
+ Package website - https://simglm.brandonlebeau.org/
+ Github - https://github.com/lebebr01/simglm
+ Twitter - https://twitter.com/blebeau11
